---
output: html_document
editor_options: 
  chunk_output_type: console
chunk_output_type: console
---
```{r}
rm(list=ls())
setwd(dirname(rstudioapi::getSourceEditorContext()$path))
```

```{r}
rm(list=ls())
dat = read.csv(file="data/clothingFullAss03.csv")
library(lme4)
fit0 <- lmer(clo~sex+(1|subjId),data=dat,REML=FALSE)

plot(fit0)
str(fit0)
```

$$
\begin{align}
Y\sim&\mathcal{N}(X\beta,Z \Psi Z+\Sigma)\\
\Sigma = &\sigma I\\
\Psi = & \sigma_B I \\
V =&Z \Psi Z+\Sigma \\
l(y,\beta,\Psi,\Sigma)\propto& -\frac{1}{2}\left\{\log|V|+(y-X\beta)^TV^{-1}(y-X\beta) \right\} \\
l_{re}(y,\beta,\Psi,\Sigma)\propto& -\frac{1}{2}\left\{\log|V|+(y-X\beta)^TV^{-1}(y-X\beta)+\log |X^T V^{-1} X | \right\}

\end{align}
$$


$$
l(y,\beta,\sigma_s,\sigma)\propto -\frac{1}{2}\sum_{i}^{N_{subjId}} \log |\sigma^2I_{n_i}+\sigma_s^2E_{{n_i},{n_i}} |+ (y_i-X_i\beta)^T\left[\sigma^2I_{n_i}+\sigma_s^2E_{{n_i},{n_i}}\right]^{-1}(y_i-X_i\beta)
$$
Where $i$ all the indices for each of the subjects, and where $n_i$ is the number of observations for each subject. $E_{n_i,n_i}$ is a matrix full of ones with the dimensions $n_i\times n_i$ and $I_{n_i}$ is the identity matrix of the same dimensions.
```{r}
pars=c("mu"=0.5,"sex_male"=-0.1,"sigma_s"=0.1,"sigma"=0.1)

kronecker(diag(3),matrix(c(1.1,1,1,1.1),ncol=2))

library(mvnfast)

X = cbind(1,rep(0,nrow(dat)))
X[dat$sex=="male",2]=1
library(Matrix)
Z = Matrix(0,nrow=nrow(dat),ncol=length(unique(dat$subjId)),sparse=T)
for(i in 1:nrow(dat)){
  Z[i,unique(dat$subjId)==dat$subjId[i]]=1
}
m = diag(c(1,2))
m[1,2]=m[2,1]=0.5
m
chol2inv(chol(m))
solve(m)
qr.solve(m)
# slow
#library(mvtnorm)


nll0<- function(pars){
  # ML estimation:
    mu = pars["mu"]
    sex_male = pars["sex_male"]
    beta=c(mu,sex_male)
    
    #y = dat$clo
    #Psi = diag(length(unique(dat$subjId)))*pars["sigma_s"]^2
    #Sigma = diag(nrow(dat))*pars["sigma"]^2
    #V = Z%*%Psi%*%t(Z)+Sigma
    # Faster dmvnorm from mvnfast
    #ll = dmvn(y,X%*%beta,V,log=T)
    
    ll=0
    
    for(i in unique(dat$subjId)){
      sex = dat$sex[dat$subjId==i][1]
      n = sum(dat$subjId==i)
      if(sex=="male"){
        x = cbind(1,rep(1,n))
      }else{
        x = cbind(1,rep(0,n))
      }
      sig = diag(n)*pars["sigma"]^2+pars["sigma_s"]^2
      y = dat$clo[dat$subjId==i]
      tmp = dmvn(y,x%*%beta,sig,log=T)
      ll = ll+tmp
    }
   
    return(-ll)
}


nll(pars)

opt0<-nlminb(pars,nll0,lower=c(-10,-10,0.001,0.001),
       upper=c(10,10,10,10))
opt0$par

fixef(fit0)
VarCorr(fit0)

logLik(fit0)
-nll0(opt0$par)



```
CALCULATE RANDOM EFFECTS??? See equation 5.60 or google

### Question 2
$u_i$ random effect for subject ID
$v_{i,j}$ random effect for interaction between subject ID and day

```{r}
library(nlme)
fit1 <- lmer(clo~sex+(1|subjId)+(1|subjId:day),data=dat,REML=F)

dat$new = paste(dat$sex,dat$day)
fit1_1 <- lme(clo~sex,data=dat,random = list(~1|subjId,~1|new),method="ML")

fit1
fit1_1
nll1<- function(pars){
  # ML estimation:
    mu = pars["mu"]
    sex_male = pars["sex_male"]
    beta=c(mu,sex_male)
    
    # for(j in unique(dat$subjId)){
    #   tmp = c()
    #   n_days = unique(dat$day[dat$subjId==j])
    #   for(k in n_days){
    #     tmp = c(tmp,sum(dat$subjId==j&dat$day==k))
    #   }
    #   cat("\n","subject",j,"\t",paste("day",n_days,"n_obs",tmp))
    # }
    # 
    
    ll=0
    for(i in unique(dat$subjId)){
      df = dat[dat$subjId==i,]
      
      sex = df$sex[1]
      n = nrow(df)
      days = unique(df$day)
      
      
      if(sex=="male"){
        x = cbind(1,rep(1,n))
      }else{
        x = cbind(1,rep(0,n))
      }
      
      sig = diag(n)*pars["sigma"]^2+pars["sigma_s"]^2
      for(k in days){
        idx = which(df$day==k)
        sig[idx,idx] = sig[idx,idx]+pars["sigma_sub_day"]^2
      }
      sig
      y = df$clo
      tmp = dmvn(y,x%*%beta,sig,log=T)
      #tmp = dmvnorm(y,x%*%beta,sig,log=T)
      #print(tmp)
      ll = ll+tmp
    }
    
    return(-ll)
}


pars=c("mu"=0.5,"sex_male"=-0.1,"sigma_sub_day"=0.1,"sigma_s"=0.1,"sigma"=0.1)

nll1(pars)

opt1<-nlminb(pars,nll1,lower=c(-10,-10,0.001,0.001,0.001),
       upper=c(10,10,10,10,10))
opt1$par
fixef(fit1)
VarCorr(fit1)

logLik(fit1)
-nll1(opt1$par)


```


### Question 3
$u_i$ random effect for interaction between subject ID and sex
$v_{i,j}$ random effect for interaction between subject ID, day and sex

```{r}

nll2<- function(pars,alpha_sex="male"){
  #print(c(pars[1:2],exp(pars[3]+pars[4:6]) ))
  #print(pars)
  # ML estimation:
  mu = pars["mu"]
  sex_male = pars["sex_male"]
  beta=c(mu,sex_male)
  alpha = pars['alpha']
  
  
  ll=0
  for(i in unique(dat$subjId)){
    #print(i)
    df = dat[dat$subjId==i,]
    df$sex
    if(alpha_sex=="male"){
      sex = df$sex_male[1]
    }
    if(alpha_sex=="female"){
      sex = df$sex_female[1]
    }
    
    n = nrow(df)
    days = unique(df$day)
    
    if(df$sex[1]=="male"){
      x = cbind(1,rep(1,n))
    }else{
      x = cbind(1,rep(0,n))
    }
    
    sig = diag(n)*exp(sex*alpha+pars["sigma^2"])+exp(sex*alpha+pars["sigma_s^2"])
    for(k in days){
      idx = which(df$day==k)
      sig[idx,idx] = sig[idx,idx]+exp(sex*alpha+pars["sigma_sub_day^2"])
    }
    
    
    y = df$clo
    tmp = dmvn(y,x%*%beta,sig,log=T)
    #tmp = dmvnorm(y,x%*%beta,sig,log=T)
    #print(tmp)
    ll = ll+tmp
  }
  return(-ll)
}
dat$sex_male = as.numeric(as.factor(dat$sex))-1
dat$sex_female = abs(as.numeric(as.factor(dat$sex))-2)
pars=c("mu"=0.5,"sex_male"=-0.1,"alpha"=1,"sigma_sub_day^2"=0.1,"sigma_s^2"=0.1,"sigma^2"=0.1)


# Fitting male alpha, female=1
opt2<-nlminb(pars,nll2,alpha_sex="male")
out_par = opt2$par
c(out_par[1:2],exp(out_par[3:6]))
exp(out_par[3])*exp(out_par[4:6])

# Fitting female alpha, male=1
opt2<-nlminb(pars,nll2,alpha_sex="female")
out_par = opt2$par
c(out_par[1:2],exp(out_par[3:6]))
exp(out_par[3])*exp(out_par[4:6])


```


```{r}

nll3<- function(pars,alpha_sex="male"){
  #print(c(pars[1:2],exp(pars[3]+pars[4:6]) ))
  #print(pars)
  # ML estimation:
  mu = pars["mu"]
  sex_male = pars["sex_male"]
  beta=c(mu,sex_male)
  alpha = pars['alpha']
  
  ll=0
  for(i in unique(dat$subjId)){
    #print(i)
    df = dat[dat$subjId==i,]
    df$sex
    if(alpha_sex=="male"){
      sex = df$sex_male[1]
    }
    if(alpha_sex=="female"){
      sex = df$sex_female[1]
    }
    
    n = nrow(df)
    days = unique(df$day)
    
    if(df$sex[1]=="male"){
      x = cbind(1,rep(1,n))
    }else{
      x = cbind(1,rep(0,n))
    }
    
    sig = diag(n)*exp(sex*alpha+pars["sigma^2"])+exp(sex*alpha+pars["sigma_s^2"])
    for(k in days){
      idx = which(df$day==k)
      sig[idx,idx] = sig[idx,idx]+exp(sex*alpha+pars["sigma_sub_day^2"])
    }
    
    
    y = df$clo
    tmp = dmvn(y,x%*%beta,sig,log=T)
    #tmp = dmvnorm(y,x%*%beta,sig,log=T)
    #print(tmp)
    ll = ll+tmp
  }
  return(-ll)
}

pars=c("mu"=0.5,"sex_male"=-0.1,"alpha"=1,"sigma_sub_day^2"=0.1,"sigma_s^2"=0.1,"sigma^2"=0.1)

# Fitting male alpha, female=1
opt2<-nlminb(pars,nll2,alpha_sex="male")
out_par = opt2$par
c(out_par[1:2],exp(out_par[3:6]))
exp(out_par[3])*exp(out_par[4:6])

# Fitting female alpha, male=1
opt2<-nlminb(pars,nll2,alpha_sex="female")
out_par = opt2$par
c(out_par[1:2],exp(out_par[3:6]))
exp(out_par[3])*exp(out_par[4:6])



len = length(unique(dat$subjId))
N = nrow(dat)
subj = unique(dat$subjId)
for(i in 1:length(subj)){
  print(paste("ID",i,"sex",dat$sex[dat$subjId==subj[i]][1],"sex_male",dat$sex_male[dat$subjId==subj[i]][1],
              "ours"))
  
}


```

## Q4

$$
\begin{align}
L(\beta,\Psi;u,y) = f_{Y|y}(y;u,\beta)f_U(u,\Psi)\\
L(\mu,\sigma,\phi;\gamma,y) = \mathcal{N}\left(\mu,\frac{\sigma}{\gamma_i}\right)G(1,\phi)\\
\\
\alpha = \phi,  \beta = \phi, x = clo\\
\\
 \frac{1}{\sqrt{2\pi}\cdot(\sigma/\gamma_i)}\exp{\left(-\frac{1}{2}\frac{(x-\mu)^2}{\sigma^2/\gamma_i^2}\right)}
\times \frac{\beta^\alpha}{\Gamma(\alpha)}\gamma^{\alpha-1}\exp{\left(-\beta \gamma\right)}\\

\frac{1}{\sqrt{2\pi}\cdot(\sigma/\gamma_i)}\exp{\left(-\frac{1}{2}\frac{(x-\mu)^2}{\sigma^2/\gamma_i^2}\right)}
\times \frac{\phi^\phi}{\Gamma(\phi)}\gamma^{\phi-1}\exp{\left(-\phi \gamma\right)}\\


\frac{\gamma \phi^\phi}{\Gamma(\phi)\times\sqrt{2\pi}\cdot(\sigma)}\exp{\left(-\phi \gamma-\frac{1}{2}\frac{\gamma^2(x-\mu)^2}{\sigma^2}\right)}
\times \gamma^{\phi-1}\\

\end{align}
$$
$$
\begin{align}
\frac{1}{\Gamma(\alpha)\beta}\left(\frac{\gamma}{\beta}\right)^{\alpha-1}\exp{\left(-\frac{\gamma}{\beta}\right)}\\

\frac{1}{\Gamma(\phi)\frac{1}{\phi}}\left(\gamma\phi \right)^{\phi-1}\exp{\left(-\gamma \phi\right)}\\
\frac{\phi}{\Gamma(\phi)}\left(\gamma \phi \right)^{\phi-1}\exp{\left(-\gamma \phi\right)} \\

\end{align}
$$

$$
\begin{align}
 \frac{\gamma}{\sqrt{2\pi}\cdot(\sigma)}\exp{\left(-\frac{1}{2}\frac{\gamma^2(x-\mu)^2}{\sigma^2}\right)} \times \frac{1}{\Gamma(\alpha)\beta}\left(\frac{\gamma}{\beta}\right)^{\alpha-1}\exp{\left(-\frac{\gamma}{\beta}\right)}\\

\frac{\gamma}{\sqrt{2\pi}\cdot \sigma}\exp{\left(-\frac{\gamma^2(x-\mu)^2}{2\sigma^2} -\frac{\gamma}{\beta}\right)} \times \frac{1}{\Gamma(\alpha)\beta}\left(\frac{\gamma}{\beta}\right)^{\alpha-1}\iff \\
\int_0^{\infty} \frac{\gamma}{\sqrt{2\pi}\cdot \sigma}\exp{\left(-\frac{\gamma^2(x-\mu)^2}{2\sigma^2} -\frac{\gamma}{\beta}\right)} \times \frac{1}{\Gamma(\alpha)\beta}\left(\frac{\gamma}{\beta}\right)^{\alpha-1} \\

\frac{1}{\sqrt{2\pi}\cdot \sigma} \frac{1}{\Gamma(\alpha)\beta} \frac{1}{\beta^{\alpha-1}}\int_0^{\infty} \gamma \exp{\left(-\frac{\gamma^2(x-\mu)^2}{2\sigma^2} -\frac{\gamma}{\beta}\right)} \times \gamma^{\alpha-1} \\

\frac{1}{\sqrt{2\pi}\cdot \sigma} \frac{1}{\Gamma(\alpha)\beta} \frac{1}{\beta^{\alpha-1}}\int_0^{\infty}  \exp{\left(-\frac{\gamma^2(x-\mu)^2}{2\sigma^2} -\frac{\gamma}{\beta}\right)} \times \gamma^{\alpha} \\

\end{align}
$$

$$
\exp (-\gamma/\beta)\gamma^{\alpha-1}
$$
gamma function
$$
\Gamma(z) = \int_0^\infty t^{z-t}\exp{(-t)} dt = 2\int_0^\infty t^{2z-t}\exp{(-t^2)} dt = \int_0^1 \left[ln\left(\frac{1}{t}\right)\right]^{z-1}dt
$$



$$
\begin{align}
pdf(x,\nu) = \frac{\Gamma(\frac{\nu+1}{2})}{\sqrt{\nu \pi}\Gamma(\frac{\nu}{2})}\left(1+\frac{x^2}{\nu}\right)^{-\frac{\nu+1}{2}}\\

\frac{1}{\sigma}pdf(\frac{y-\mu}{\sigma},2\phi) = \sigma \frac{\Gamma(\frac{2\phi+1}{2})}{\sqrt{2\phi \pi}\Gamma(\frac{2\phi}{2})}\left(1+\frac{x^2}{2\phi}\right)^{-\frac{2\phi+1}{2}} \\

\frac{1}{\sigma}pdf(\frac{y-\mu}{\sigma},2\phi) = \sigma \frac{\Gamma(\frac{2\phi+1}{2})}{\sqrt{2\phi \pi}\Gamma(\phi)}\left(1+\frac{(\frac{y-\mu}{\sigma})^2}{2\phi}\right)^{-\frac{2\phi+1}{2}} 
\end{align}

$$
### WORKS!!

Assuming: $ Y_i|\gamma_i \sim \mathcal{N}(\mu,\sigma^2/\gamma_i)$
https://people.eecs.berkeley.edu/~jordan/courses/260-spring10/lectures/lecture5.pdf

$$
\begin{align}
\int_0^\infty \frac{\sqrt{\gamma}}{\sqrt{2\pi}\cdot(\sigma)}\exp{\left(-\frac{1}{2}\frac{\gamma(x-\mu)^2}{\sigma^2}\right)} 
\times \frac{\beta^\alpha}{\Gamma(\alpha)}\gamma^{\alpha-1}\exp{(-\gamma\beta)}\\

= \frac{1}{\sqrt{(2\pi)}\sigma} \frac{\beta^\alpha}{\Gamma(\alpha)} \int_0^\infty \sqrt{\gamma}\exp{\left(-\frac{1}{2}\frac{\gamma(x-\mu)^2}{\sigma^2}\right)}
\times\gamma^{\alpha-1}\exp{(-\gamma\beta)} \\

= \frac{1}{\sqrt{(2\pi)}\sigma} \frac{\beta^\alpha}{\Gamma(\alpha)} \int_0^\infty \gamma^{(\alpha+\frac{1}{2})-1}\exp{\left(-\gamma\beta-\frac{1}{2}\frac{\gamma(x-\mu)^2}{\sigma^2}\right)}
 \\
 
 = \frac{1}{\sqrt{(2\pi)}\sigma} \frac{\beta^\alpha}{\Gamma(\alpha)} \int_0^\infty \gamma^{(\alpha+\frac{1}{2})-1}\exp{\left(-\gamma \left(\beta+\frac{1}{2}\frac{(x-\mu)^2}{\sigma^2}\right)\right)}
 \\
 
 = \frac{1}{\sqrt{(2\pi)}\sigma} \frac{\beta^\alpha}{\Gamma(\alpha)} \int_0^\infty \gamma^{\hat{\alpha}-1}\exp{\left(-\gamma \hat{\beta}\right)}, \hspace{0.5cm}\hat{\beta}=\left(\beta+\frac{1}{2}\frac{(x-\mu)^2}{\sigma^2}\right), \hat{\alpha}=\alpha+1/2\\
 
= \frac{1}{\sqrt{(2\pi)}\sigma} \frac{\beta^\alpha}{\Gamma(\alpha)} \frac{\Gamma(\hat{\alpha})}{\hat{\beta}^{\hat{\alpha}}}\\
 
 
= \frac{1}{\sqrt{(2\pi)}\sigma} \frac{\beta^\alpha}{\Gamma(\alpha)} \frac{\Gamma(\alpha+1/2)}{\left(\beta+\frac{1}{2}\frac{(x-\mu)^2}{\sigma^2}\right)^{\alpha+1/2}}\\
 
\end{align}
$$

$$ 
\begin{align}
\frac{1}{\sigma}\frac{\Gamma(\phi+1/2)}{\sqrt{2\pi}\Gamma(\phi)}\phi^\phi
\left(\phi+\frac{(x-\mu)^2}{2\sigma^2}\right)^{-\phi-1/2}\\

\frac{1}{\sigma}\frac{\Gamma(\phi+1/2)}{\sqrt{2\pi}\Gamma(\phi)}\phi^\phi
\left(1+\frac{(x-\mu)^2}{2\sigma^2\phi}\right)^{-\phi-1/2}\cdot \left(\frac{1}{\phi}\right)^{-\phi-1/2}\\


\frac{1}{\sigma}\frac{\Gamma(\phi+1/2)}{\sqrt{2\pi}\Gamma(\phi)}
\left(1+\frac{(x-\mu)^2}{2\sigma^2\phi}\right)^{-\phi-1/2}\cdot \left(\frac{1}{\phi}\right)^{-1/2}\\

\frac{1}{\sigma}\frac{\Gamma(\phi+1/2)}{\sqrt{2\pi \phi}\Gamma(\phi)}
\left(1+\frac{(x-\mu)^2}{2\sigma^2\phi}\right)^{-\phi-1/2}\\

\end{align}
$$

$$\frac{1}{\sigma}pdf(\frac{y-\mu}{\sigma},2\phi) = \frac{1}{\sigma} \frac{\Gamma(\phi+\frac{1}{2})}{\sqrt{2\phi \pi}\Gamma(\phi)}\left(1+\frac{(\frac{y-\mu}{\sigma})^2}{2\phi}\right)^{-\phi-\frac{1}{2}} $$







