---
output: html_document
editor_options: 
  chunk_output_type: console
chunk_output_type: console
---

```{r}
source('setup.R')
```

```{r}
rm(list=ls())
library(ggplot2)
library(knitr)
library(RColorBrewer)
library(tidyverse)
library(latex2exp)
library(car)
library(xtable)
library(gridExtra)
library(nlme)
```

# Part 1 Clothing insulation level
```{r}
# Loading the data
df = read.csv(file="data/clothingFullAss03.csv")
df$time2=factor(df$time2)
df$day = factor(df$day)
df$subjId = factor(df$subjId)
df$sex = factor(df$sex)
```

## Exploratory data analysis

```{r part1_exploratory,fig.cap="This graph shows the Clothing level as a function Indoor Operating temperatures. This is further split into day and gender, and in each subplot the observations for each person are connected."}
# Exploratory analysis
p1=ggplot(df,aes(x=tOut,y=clo,col=subjId))+
  geom_path()+
  geom_text(aes(label=time2))+
  facet_grid(rows=vars(sex),cols=vars(paste("day",day)))+
  theme(legend.position = "None")+
  ggtitle("Clothing Level vs Outdoor Temperature")+
  xlab("Outdoor Temperature")+ylab("Clothing level")

p2=ggplot(df,aes(x=tInOp,y=clo,col=subjId))+
  geom_path()+
  geom_text(aes(label=time2))+
  facet_grid(rows=vars(sex),cols=vars(paste("day",day)))+
  theme(legend.position = "None")+
  ggtitle("Clothing Level vs Indoor Operating Temperature")+
  xlab("Indoor Operating Temperature")+ylab("Clothing level")

#grid.arrange(p1,p2,nrow=2)
p2

```

From figure \ref{fig:part1_exploratory} it appears that females have a larger variation in clothing level compared to males. We see that some subjects have different preferred clothing level and this seems especially to be the case for females. It should also be noted that this figure also shows the auto-correlation of each subject within each day and there appears to be correlation within each day. This is seen as clothing level appears to be more dependent on the subject within the given day rather than between the days. We do not see an increase/decrease in clothing level for each day which is why we will consider the variable as a factor. 

Further exploratory analysis will not be performed as this already have been presented in assignment 1. 


## Initial fixed effect models with subject ID as random effect
We will fit a mixed effect model with subject ID as a random effect. The initial model will predict clothing level based on gender of the subject ($sex$), Outdoor temperature ($tOut$), Indoor Operating temperature ($tInOp$) and time ($time$). The initial model is:
$$ clo_i = \beta_0  + \beta_1sex_i +  \beta_2tOut_i +  \beta_3 tInOp_i+ \beta_{4t}time_{it} + \beta_5sex_i(tOut_i) + \beta_6sex_i(tInOp_i) + \beta_{7d} day_{id}+ SubjId_i + \epsilon_{ij}$$
where $SUbjId_i \sim N(0,\sigma_u^2)$, $\epsilon_{ij} \sim N(0,\sigma^2)$ and both are mutually independent. $sex_i(tOut_i)$ indicates the interaction term between gender of the subject and the Outdoor temperature. The model will be reduced based on backward selection using Type III partitioning where higher order terms are removed first. Firstly, it is determined whether time should be included as a factor. Two models are generated according to above equation where the only difference are whether time are included as a factor or numeric. The log-likelihood of the two models are: 706.5 and 711.6 for the model including time as numeric and factor respectively and thereby indicating that time should be included as a factor.

```{r}
#We will not estimate individual slopes as the clothing levels seem to be constant for each subject during each day. WE PROBABLY SHOULD ANYWAYS!!! TRY RANDOM SLOPES

init_fit2 = lme(clo~sex*(tOut+tInOp)+time,random=~ 1|subjId,data=df,method="ML")
init_fit4 = lme(clo~sex*(tOut+tInOp)+time2,random=~ 1|subjId,data=df,method="ML")


anova(init_fit2,init_fit4)

```

The model are reduced based on maximum likelihood estimation and the final model are: 
INDSÆT TABEL OVER HVAD DER FJERNES OG HVORNÅR 

```{r}
fit0_2 = lme(clo~sex*(tOut+tInOp)+day+time2,random=~ 1|subjId,data=df,method="ML")
Anova(fit0_2,type=3)

fit1_2 = lme(clo~sex*tInOp+tOut+day+time2,random=~ 1|subjId,data=df,method="ML")
Anova(fit1_2,type=3)

anova(fit0_2, fit1_2)
```

$$ clo_i = \beta_0  + \beta_1sex_i +  \beta_2tOut_i +  \beta_3 tInOp_i+ \beta_{4t}time_{it} + \beta_5sex_i(tInOp_i) + \beta_{6d} day_{id}+ SubjId_i + \epsilon_{ij}$$
A likelihood ratio test are performed to test the model reduction and results in $p = 0.52$ and the model reduction is therefore kept.

SKAL VI TESTE VORES RANDOM EFFECT PARAMETER OM DEN SKAL INKLUDERES VED LIKELIHOOD RATIO TEST?

Afterwards the random effect coefficients are estimated using REML estimation. The coefficients for the random effects are presented in the qq-plot in figure \ref{fig:hist_1} and does appear to be randomly distributed as assumed in the model.The fixed effect coefficients are presented in table \ref{tab:coef_1} including 95 \% confidence intervals.

```{r hist_1,fig.cap="This graph shows the random effect coefficients presented in a QQ-plot"}
fit1_2 = lme(clo~sex*tInOp+tOut+day+time2,random=~ 1|subjId,data=df,method="REML")
qqnorm(ranef(fit1_2)[,1], xlab = 'Random effect coefficients', main = 'Histogram of random coefficients')
qqline(ranef(fit1_2)[,1], xlab = 'Random effect coefficients', main = 'Histogram of random coefficients')

xtable(intervals(fit1_2)$fixed)
```

\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & Lower (2.5 \%)& Estimate & Upper (97.5 \%)\\ 
  \hline
(Intercept) & 0.98 & 1.17 & 1.36 \\ 
  sexmale & -0.66 & -0.42 & -0.17 \\ 
  tInOp & -0.02 & -0.01 & -0.01 \\ 
  tOut & -0.01 & -0.01 & -0.01 \\ 
  day2 & -0.02 & -0.01 & 0.01 \\ 
  day3 & -0.05 & -0.03 & -0.02 \\ 
  day4 & 0.08 & 0.15 & 0.22 \\ 
  time22 & 0.02 & 0.04 & 0.06 \\ 
  time23 & 0.03 & 0.05 & 0.08 \\ 
  time24 & 0.03 & 0.05 & 0.08 \\ 
  time25 & 0.03 & 0.06 & 0.08 \\ 
  time26 & 0.03 & 0.06 & 0.09 \\ 
  sexmale:tInOp & 0.00 & 0.01 & 0.02 \\ 
   \hline
\end{tabular}
\caption{Estimate of fixed effect coefficients including 95\% confidence intervals.}
\label{tab:coef_1}
\end{table}

The residuals are further investigated in figure \ref{fig:res_1} where the standardized residuals are plotted against the fitted values. From this the residuals appears to be randomly distributed.

```{r res_1,fig.cap="This graph shows the residuals of the first fixed effect model with subject Id as random effect."}
plot(fit1_2)
```


## Fixed effect models with subject ID and day as random effects

A new fixed effect model with subject ID and day as random effects will be used for modelling the clothing insulation level. The initial model has the following set-up:
$$ clo_i = \beta_0  + \beta_1sex_i +  \beta_2tOut_i +  \beta_3 tInOp_i+ \beta_{4t}time_{it} + \beta_5sex_i(tInOp_i) + SubjId_i +day_{id} +  \epsilon_{ij}$$
The random effect day is nested within the random effect Subject ID. Here $SUbjId_i \sim N(0,\sigma_u^2)$, $\epsilon_{ij} \sim N(0,\sigma^2)$ and $day_{id} \sim N(0, \sigma_d^2)$ and all three are mutually independent. ER DAY AFHÆNGIG AF SUBJECT ID??. Again $sex_i(tOut_i)$ indicates the interaction term between gender of the subject and the Outdoor temperature. Again time will be included as a factor based on the result in the previous subsection.

The initial model will be reduced using backward selection based on type III partitioning where higher order terms are removed first based on maximum-likelihood estimation. INDSÆT TABEL FOR MODEL REDUCTION 

```{r}
fit0 = lme(clo~sex*(tOut+tInOp)+time2,random=list(~1|subjId,~1|day),data=df,method="ML")
Anova(fit0,type=3)
fit1 = lme(clo~sex*tInOp+tOut+time2,random=list(~1|subjId,~1|day),data=df,method="ML")
Anova(fit1,type=3)

anova(fit0, fit1)
```
The final model after model reduction are: 
$$ clo_i = \beta_0  + \beta_1sex_i +  \beta_2tOut_i +  \beta_3 tInOp_i+ \beta_{4t}time_{it} + \beta_5sex_i(tInOp_i) + SubjId_i +day_{id} +  \epsilon_{ij}$$

A likelihood ratio test are performed to test the model reduction and results in $p = 0.48$ and the model reduction is therefore kept.


The fixed effect coefficients are presented in table \ref{tab:coef_2} including 95 \% confidence intervals. It should be noted that the majority of the contribution comes from the gender of the subject ie. the intercept and the sex coefficient. In figure \ref{fig:hist_2} that the day random effects appear to be normal-distributed with some tails. The subjectId random effects appears to have some heavy tails, and this might indicate that modeling these as normal distributed might not be the correct choice. 

```{r hist_2,fig.cap="This graph shows the random effect coefficients presented in a QQ-plot"}
xtable(intervals(fit1)$fixed)
fit1 = lme(clo~sex*tInOp+tOut+time2,random=list(~1|subjId,~1|day),data=df,method="REML")
par(mfrow=c(1,2))
qqnorm(ranef(fit1)$day[,1],xlab = 'Random effect coefficients (day)', main = 'QQ-plot of random coefficients for day')
qqline(ranef(fit1)$day[,1])
qqnorm(ranef(fit1)$subjId[,1],xlab = 'Random effect coefficients (subjectID)', main = 'QQ-plot of random coefficients for subjectID')
qqline(ranef(fit1)$subjId[,1])

```

\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & Lower (2.5 \%)& Estimate & Upper (97.5 \%)\\ 
  \hline
(Intercept) & 0.85 & 1.00 & 1.15 \\ 
  sexmale & -0.54 & -0.37 & -0.21 \\ 
  tInOp & -0.02 & -0.01 & -0.01 \\ 
  tOut & -0.01 & -0.00 & -0.00 \\ 
  time22 & 0.01 & 0.03 & 0.04 \\ 
  time23 & 0.01 & 0.03 & 0.05 \\ 
  time24 & 0.00 & 0.02 & 0.04 \\ 
  time25 & 0.00 & 0.02 & 0.04 \\ 
  time26 & 0.00 & 0.03 & 0.05 \\ 
  sexmale:tInOp & 0.01 & 0.01 & 0.02 \\ 
   \hline
\end{tabular}
\caption{Estimate of fixed effect coefficients including 95\% confidence intervals.}
\label{tab:coef_2}

\end{table}


The residuals are further investigated in figure \ref{fig:res_2} where the standardized residuals are plotted against the fitted values. From this the residuals appears to contain some structure going from small to large residuals when increasing the fitted values. HVAD BETYDER DET?? Autocorrelation??

```{r res_2,fig.cap="This graph shows the residuals of the first fixed effect model with subject Id and day as random effect."}
plot(fit1)
```



## Mixed effect model with autocorrelation
Now, a mixed effect model with within day autocorrelation will be tested. The initial model has the following set-up

$$
clo_i = \beta_0  + \beta_1sex_i +  \beta_2tOut_i +  \beta_3 tInOp_i+ \beta_4sex_i(tInOp_i) + \beta_5sex_i(tOut_i) + subDay_s +  \epsilon_{ij}
$$
where different autocorrelation-structures for time are tested with subday as a random effect. First, the auto-correlation structure will be investigated where a auto-regressive(1), gaussian and exponential correlation structure will be tested. It should be noted, that for the gaussian and exponential correlations structure time isn't required to be equidistant. For AR(1) time has to be equidistant however. 

```{r}
df$day = as.numeric(as.character(df$day))
init_fit1_ar = lme(clo~sex*(tOut+tInOp),random=~1|subDay,
                correlation = corAR1(form=~as.numeric(as.character(time2))|subDay),
                data=df,method="ML")


fit_gaus = lme(clo~sex*(tInOp+tOut),random=~1|subDay,
                correlation = corGaus(form=~as.numeric(as.character(time))|subDay),
                data=df,method="ML")

fit_gaus_1 = lme(clo~sex*(tInOp+tOut),random=~1|subDay,
                correlation = corGaus(form=~as.numeric(as.character(time))|subDay, nugget = T),
                data=df,method="ML")

fit_exp_1 = lme(clo~sex*(tInOp+tOut),random=~1|subDay,
                correlation = corExp(form=~as.numeric(as.character(time))|subDay),
                data=df,method="ML")

xtable(anova(fit_gaus,fit_gaus_1, init_fit1_ar,fit_exp_1))


```

\begin{table}[ht]
\centering
\begin{tabular}{rlrrrrrlrr}
  \hline
 & call & Model & df & AIC & BIC & logLik\\ 
  \hline
  Gaussian & corGaus(\~time $|$subDay), &   1 & 9.00 & -1992.40 & -1950.20 & 1005.20  \\ 
  Gaussian w/ nugget & corGaus(~time $|$subDay, nugget = T) &   2 & 10.00 & -2008.08 & -1961.20 & 1014.04 \\ 
  AR(1) & corAR1(\~time2 $|$ subDay) &   3 & 9.00 & -2019.59 & -1977.39 & 1018.79  \\ 
  Exponential & corExp(\~time $|$ subDay) &   4 & 9.00 & -2011.36 & -1969.17 & 1014.68  \\ 
   \hline
\end{tabular}
\label{tab:1.51}
\end{table}

As can be seen in table \ref{tab:q1.5}, all the information criteria points to using a auto-regressive(1) correlation structure. The model will be reduced using type 3 partitioning and backwards selection. Higher order terms will be removed before lower order terms. 



```{r}

init_fit1 = lme(clo~sex*(tOut+tInOp),random=~1|subDay,
                correlation = corAR1(form=~as.numeric(as.character(time2))|subDay),
                data=df,method="REML")

Anova(init_fit1,type=3)

init_fit2 = lme(clo~tOut+sex*tInOp,random=~1|subDay,
                correlation = corAR1(form=~as.numeric(as.character(time2))|subDay),
                data=df,method="REML")

Anova(init_fit2,type=3)


init_fit1_ml = lme(clo~sex*(tOut+tInOp),random=~1|subDay,
                correlation = corAR1(form=~as.numeric(as.character(time2))|subDay),
                data=df,method="ML")


init_fit2_ml = lme(clo~tOut+sex*tInOp,random=~1|subDay,
                correlation = corAR1(form=~as.numeric(as.character(time2))|subDay),
                data=df,method="ML")

anova(init_fit1_ml, init_fit2_ml)
init_fit2_ml = lme(clo~tOut+sex*tInOp,random=~1|subDay,
                correlation = corAR1(form=~as.numeric(as.character(time2))|subDay),
                data=df,method="REML")

xtable(intervals(init_fit2)$fixed)

```

Only the interactions between sex and outdoor temperature is removed. The final model formulation is then given by

$$
clo_i = \beta_0  +  \beta_1tOut_i +  \beta_2sex_i + \beta_3 tInOp_i+  \beta_4sex_i(tInOp_i) + subDay_s +  \epsilon_{ij}.
$$
A likelihood ratio test is performed to see if the model reduction is significant. This results in a p-value of 0.44 and the model reduction is thereby kept. In table \ref{tab:q1_52} the estimated parameters and their 95 \% confidence interval is given. Additionally, the AR(1) coefficient $\phi$ was estimated to 0.56. 

\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & Lower (2.5 \%)& Estimate & Upper (97.5 \%)\\ 
  \hline
(Intercept) & 0.76 & 0.90 & 1.03 \\ 
  tOut & -0.01 & -0.00 & -0.00 \\ 
  sexmale & -0.54 & -0.34 & -0.15 \\ 
  tInOp & -0.01 & -0.01 & -0.00 \\ 
  sexmale:tInOp & 0.00 & 0.01 & 0.02 \\ 
   \hline
\end{tabular}
\label{tab:q1_52}
\end{table}



## Interpretation of model
The model intercept is the female contribution as sex(female) = 0 in our model formulation. Sexmale is the correction to the intercept for male subjects. The same applies for the indoor operating temperature, where it should be noted that the male contribution almost cancels out the effect of change in indoor operating temperature. This could indicate the males are not affected in their clothing level by the indoor operating temperature. The numeric largest fixed coefficient is the gender coefficient which implies it has a big impact on clothing level compared to the other fixed coefficients. 

The random effect will now be investigated. 

```{r}
qqnorm(ranef(init_fit2)[,1])
qqline(ranef(init_fit2)[,1])

```

Here, it the heavy tails should be noted. They might indicate that another distribution for the random effect should be applied where the heavy tails are accounted for. 

```{r variogram, fig.cap="Variograms for Gaussian and Exponential correlation structure."}
par(mfrow = c(2,1))
plot(Variogram(fit_gaus_1), main = "Gaussian")
plot(Variogram(fit_exp_1), main = "Exponential")
```

From figure \ref{fig:variogram}, both the Gaussian and Exponential correlation structure seem to fit well. As AR(1) had the highest likelihood and a AR(1) correlation structure resembles the Exponential correlation structure, it is assumed that the AR(1) is also a good choice for the correlation structure. 


```{r res_analysis, fig.cap="Residual analysis of final model. The first plot is the qqplot of the residuals split for each gender. The two bottom plots are the residuals dependent on the outdoor temperature also split for each gender. Additionaly, the residuals are grouped accordingly to the random effect subday and colored according to subjectId."}

library(gridExtra)
df$residuals <- resid(init_fit2)

p1 <- ggplot(data = df, aes(sample = residuals, col = sex))+
  stat_qq() + stat_qq_line()


p2 <- ggplot(data = df, aes(x = tOut, y = residuals, col = subjId, group = subDay)) +
  geom_line()+
  geom_point()+
  facet_wrap(~sex)+
  theme(legend.position = "None")

grid.arrange(p1,p2)
# Might be better  fits for tails
```

From the qqplot in figure \ref{fig:res_analysis}, it should be noted that the model captures the data for males well with only one subject giving tails. This is not the case for females, however, as can also be seen in the qqplot. Here, heavy tails are present at both sides which indicates that there is still some variance in the data for females that is not accounted for. The same applies to the two bottom plots in figure \ref{fig:res_analysis}. The residuals for females dependent on outdoor temperature still seem randomly distributed. This could indicate that another distribution for the random effects than currently assumed would model the data better. Introducing an overdispersion parameter would maybe account for the higher variance in the female group as the residuals appears to be randomly distributed.


## Conclusion
```{r}
fit1 = lme(clo~sex*tInOp+tOut+day+time2,random=~ 1|subjId,data=df,method="REML")
fit2 = lme(clo~sex*tInOp+tOut+time2,random=list(~1|subjId,~1|day),data=df,method="REML")
fit3 = lme(clo~tOut+sex*tInOp,random=~1|subDay,
                correlation = corAR1(form=~as.numeric(as.character(time2))|subDay),
                data=df,method="REML")

AIC(fit1,fit2, fit3)
```
Three different models were proposed using different set-up of the random effects and correlation structure. The complexity of the models are increased in order to model the clothing level. 
The first model introduced had the following setup: 

$$ clo_i = \beta_0  + \beta_1sex_i +  \beta_2tOut_i +  \beta_3 tInOp_i+ \beta_{4t}time_{it} + \beta_5sex_i(tInOp_i) + \beta_{6d} day_{id}+ SubjId_i + \epsilon_{ij}$$

Where subject ID were used as random effect, this resulted in AIC -1326.5 with 15 degrees of freedom. To further improve this model day was introduced as a nested random effect of the subject ID. The used resulting model structure was: 

$$ clo_i = \beta_0  + \beta_1sex_i +  \beta_2tOut_i +  \beta_3 tInOp_i+ \beta_{4t}time_{it} + \beta_5sex_i(tInOp_i) + SubjId_i +day_{id} +  \epsilon_{ij}$$

By introducing day as a nested random effect this improved the model performance resulting in AIC -1837.2 and 13 degrees of freedom. Finally, due to the repeated measurements a AR(1) correlation structure of time given the subDay feature was implemented. The model set-up was the following:

$$
clo_i = \beta_0  +  \beta_1tOut_i +  \beta_2sex_i + \beta_3 tInOp_i+  \beta_4sex_i(tInOp_i) + subDay_s +  \epsilon_{ij}.
$$

This resulted in AIC of -1976.7 with 8 degrees of freedom. It is therefore apparent that by modeling more complex structure leads to improved model performance at the cost of a few degrees of freedom. The random effects of the final model did not appear to be normal distributed (at least for females) which could indicate that these should not be modeled as normal distributed but instead using distributions with heavier tails such as the t-distribution. 




```{r}
#Exercise 2: 
#We will not estimate individual slopes as the clothing levels seem to be constant for each subject during each day. WE PROBABLY SHOULD ANYWAYS!!! TRY RANDOM SLOPES

# Try 1+tOut|subjId or 1+tInOp|subjId 

# Does not make sense
#init_fit1 = lme(clo~sex*(tOut+tInOp)+time,random=~ 1+sex|subjId,data=df,method="ML")
# init_fit2 = lme(clo~sex*(tOut+tInOp)+time,random=~ 1|subjId,data=df,method="ML")
#init_fit3 = lme(clo~sex*(tOut+tInOp)+time2,random=~ 1+sex|subjId,data=df,method="ML")
# init_fit4 = lme(clo~sex*(tOut+tInOp)+time2,random=~ 1|subjId,data=df,method="ML")
# 
# 
# 
# ranef(init_fit1)[,1]
# hist(ranef(init_fit1)[,1])
# hist(ranef(init_fit1)[,2])
# 
# df2=data.frame(ranef1 =ranef(init_fit1)[,1],ranef2 = ranef(init_fit1)[,2],sex =unique(df[,c("sex","subjId")])$sex  )
# df2$sum = df2$ranef1
# df2$sum[df2$sex=="male"]=df2$sum[df2$sex=="male"]+df2$ranef2[df2$sex=="male"]
# 
# p1=ggplot(df2,aes(x=sum,col=sex))+geom_histogram()+facet_wrap(~sex)
# 
# 
# df3=data.frame(intercept = ranef(init_fit2)[,1],sex =unique(df[,c("sex","subjId")])$sex )
# df3
# p2=ggplot(df3,aes(x=intercept,col=sex))+geom_histogram()+facet_wrap(~sex)
# grid.arrange(p1,p2,nrow=2)
# 
# rbind(df3$intercept,df2$sum)
# 
# anova(init_fit1,init_fit2,init_fit3,init_fit4)
# unique(df$subjId)
# small difference between the intercepts
# We will continue with both models
#test = lmer(clo~sex*(tOut+tInOp)+day+time2+(1+sex|subjId),data=df)

# 
# fit0_1 = lme(clo~sex*(tOut+tInOp)+day+time2,random=~ 1+sex|subjId,data=df,method="REML")
# fit0_2 = lme(clo~sex*(tOut+tInOp)+day+time2,random=~ 1|subjId,data=df,method="REML")

# #drop1(fit0_1,test="Chisq")
# # Anova(fit0_1,type=3)
# #drop1(fit0_2,test="Chisq")
# Anova(fit0_2,type=3)
# anova(fit0_1,fit0_2)
# 
# fit1_1 = lme(clo~sex*tInOp+tOut+day+time2,random=~ 1+sex|subjId,data=df,method="REML")
# fit1_2 = lme(clo~sex*tInOp+tOut+day+time2,random=~ 1|subjId,data=df,method="REML")
# #drop1(fit1_1,test="Chisq")
# #drop1(fit1_2,test="Chisq")
# Anova(fit1_1,type=3)
# Anova(fit1_2,type=3)
# anova(fit1_1,fit1_2)
# intervals(fit1_1,which="fixed")
# intervals(fit1_1,which="var-cov")
# confint(ranef(fit1_1))
# intervals(fit1_2)
# 
# plot(fit1_1)
# plot(fit1_2)
# ```

```{r}
# #Task 2.3
# init_fit2 = lme(clo~sex*(tOut+tInOp)+time,random=list(~1|subjId,~1|day),data=df,method="ML")
# init_fit2
# 
# init_fit4 = lme(clo~sex*(tOut+tInOp)+time2,random=list(~1|subjId,~1|day),data=df,method="ML")
# anova(init_fit1,init_fit2,init_fit3,init_fit4)
# 
# 
# #fit0_1 = lme(clo~sex*(tOut+tInOp)+time2,random=list(~1+sex|subjId,~1+sex|day),data=df,method="REML")
# fit0_2 = lme(clo~sex*(tOut+tInOp)+time2,random=list(~1|subjId,~1|day),data=df,method="REML")
# anova(fit0_1,fit0_2)
# Anova(fit0_1,type=3)
# Anova(fit0_2,type=3)
# 
# 
# #fit1_1 = lme(clo~sex*tInOp+tOut+time2,random=list(~1+sex|subjId,~1+sex|day),data=df,method="REML")
# fit1_2 = lme(clo~sex*tInOp+tOut+time2,random=list(~1|subjId,~1|day),data=df,method="REML")
# anova(fit1_1,fit1_2)
# Anova(fit1_1,type=3)
# Anova(fit1_2,type=3)
# 
# intervals(fit1_1)
# 
# intervals(fit1_2)


```

