---
output: html_document
editor_options: 
  chunk_output_type: console
chunk_output_type: console
---
```{r}
setwd(dirname(rstudioapi::getSourceEditorContext()$path))
```

```{r}
rm(list=ls())
dat = read.csv(file="data/clothingFullAss03.csv")
library(lme4)
fit0 <- lmer(clo~sex+(1|subjId),data=dat,REML=FALSE)

plot(fit0)
str(fit0)
```

$$
\begin{align}
Y\sim&\mathcal{N}(X\beta,Z \Psi Z+\Sigma)\\
\Sigma = &\sigma I\\
\Psi = & \sigma_B I \\
V =&Z \Psi Z+\Sigma \\
l(y,\beta,\Psi,\Sigma)\propto& -\frac{1}{2}\left\{\log|V|+(y-X\beta)^TV^{-1}(y-X\beta) \right\} \\
l_{re}(y,\beta,\Psi,\Sigma)\propto& -\frac{1}{2}\left\{\log|V|+(y-X\beta)^TV^{-1}(y-X\beta)+\log |X^T V^{-1} X | \right\}

\end{align}
$$


$$
l(y,\beta,\sigma_s,\sigma)\propto -\frac{1}{2}\sum_{i}^{N_{subjId}} \log |\sigma^2I_{n_i}+\sigma_s^2E_{{n_i},{n_i}} |+ (y_i-X_i\beta)^T\left[\sigma^2I_{n_i}+\sigma_s^2E_{{n_i},{n_i}}\right]^{-1}(y_i-X_i\beta)
$$
Where $i$ all the indices for each of the subjects, and where $n_i$ is the number of observations for each subject. $E_{n_i,n_i}$ is a matrix full of ones with the dimensions $n_i\times n_i$ and $I_{n_i}$ is the identity matrix of the same dimensions.
```{r}
pars=c("mu"=0.5,"sex_male"=-0.1,"sigma_s"=0.1,"sigma"=0.1)

library(mvnfast)

X = cbind(1,rep(0,nrow(dat)))
X[dat$sex=="male",2]=1
library(Matrix)
Z = Matrix(0,nrow=nrow(dat),ncol=length(unique(dat$subjId)),sparse=T)
for(i in 1:nrow(dat)){
  Z[i,unique(dat$subjId)==dat$subjId[i]]=1
}
m = diag(c(1,2))
m[1,2]=m[2,1]=0.5
m
chol2inv(chol(m))
solve(m)
qr.solve(m)
# slow
library(mvtnorm)


nll0<- function(pars){
  # ML estimation:
    mu = pars["mu"]
    sex_male = pars["sex_male"]
    beta=c(mu,sex_male)
    
    #y = dat$clo
    #Psi = diag(length(unique(dat$subjId)))*pars["sigma_s"]^2
    #Sigma = diag(nrow(dat))*pars["sigma"]^2
    #V = Z%*%Psi%*%t(Z)+Sigma
    # Faster dmvnorm from mvnfast
    #ll = dmvn(y,X%*%beta,V,log=T)
    
    test=0
    for(i in unique(dat$subjId)){
      sex = dat$sex[dat$subjId==i][1]
      n = sum(dat$subjId==i)
      if(sex=="male"){
        x = cbind(1,rep(1,n))
      }else{
        x = cbind(1,rep(0,n))
      }
      sig = diag(n)*pars["sigma"]^2+pars["sigma_s"]^2
      y = dat$clo[dat$subjId==i]
      tmp = dmvn(y,x%*%beta,sig,log=T)
      test = test+tmp
    }
    ll = test
    
    return(-ll)
}


nll(pars)

opt0<-nlminb(pars,nll0,lower=c(-10,-10,0.001,0.001),
       upper=c(10,10,10,10))
opt0$par

fixef(fit0)
VarCorr(fit0)

logLik(fit0)
-nll(opt0$par)


```


### Question 2
$u_i$ random effect for subject ID
$v_{i,j}$ random effect for interaction between subject ID and day

```{r}
fit1 <- lmer(clo~sex+(1|subjId)+(1|subjId:day),data=dat,REML=F)

nll1<- function(pars){
  # ML estimation:
    mu = pars["mu"]
    sex_male = pars["sex_male"]
    beta=c(mu,sex_male)
    
    # for(j in unique(dat$subjId)){
    #   tmp = c()
    #   n_days = unique(dat$day[dat$subjId==j])
    #   for(k in n_days){
    #     tmp = c(tmp,sum(dat$subjId==j&dat$day==k))
    #   }
    #   cat("\n","subject",j,"\t",paste("day",n_days,"n_obs",tmp))
    # }
    # 
    
    ll=0
    for(i in unique(dat$subjId)){
      df = dat[dat$subjId==i,]
      
      sex = df$sex[1]
      n = nrow(df)
      days = unique(df$day)
      
      
      if(sex=="male"){
        x = cbind(1,rep(1,n))
      }else{
        x = cbind(1,rep(0,n))
      }
      
      sig = diag(n)*pars["sigma"]^2+pars["sigma_s"]^2
      for(k in days){
        idx = which(df$day==k)
        sig[idx,idx] = sig[idx,idx]+pars["sigma_sub_day"]^2
      }
      sig
      y = df$clo
      tmp = dmvn(y,x%*%beta,sig,log=T)
      #tmp = dmvnorm(y,x%*%beta,sig,log=T)
      #print(tmp)
      ll = ll+tmp
    }
    
    return(-ll)
}


pars=c("mu"=0.5,"sex_male"=-0.1,"sigma_sub_day"=0.1,"sigma_s"=0.1,"sigma"=0.1)

nll1(pars)

opt1<-nlminb(pars,nll1,lower=c(-10,-10,0.001,0.001,0.001),
       upper=c(10,10,10,10,10))
opt1$par
fixef(fit1)
VarCorr(fit1)

logLik(fit1)
-nll1(opt1$par)


```


### Question 3
$u_i$ random effect for interaction between subject ID and sex
$v_{i,j}$ random effect for interaction between subject ID, day and sex

```{r}

nll2<- function(pars,alpha_sex="male"){
  # ML estimation:
    mu = pars["mu"]
    sex_male = pars["sex_male"]
    beta=c(mu,sex_male)
    if(alpha_sex=="male"){
      alpha_male = exp(pars["alpha_male"])
    }else{
      alpha_female = exp(pars["alpha_female"])
    }
    
    
    ll=0
    for(i in unique(dat$subjId)){
      df = dat[dat$subjId==i,]
      sex = df$sex[1]
      n = nrow(df)
      days = unique(df$day)
      if(sex=="male"){
        x = cbind(1,rep(1,n))
        if(alpha_sex=="male"){
          alpha = alpha_male
        }else{
          alpha=1
        }
        
      }else{
        x = cbind(1,rep(0,n))
        if(alpha_sex=="female"){
          alpha = alpha_female
        }else{
          alpha=1
        }
      }
    
      sig = diag(n)*alpha*exp(pars["sigma^2"])+alpha*exp(pars["sigma_s^2"])
      for(k in days){
        idx = which(df$day==k)
        sig[idx,idx] = sig[idx,idx]+alpha*exp(pars["sigma_sub_day^2"])
      }
      sig
      y = df$clo
      tmp = dmvn(y,x%*%beta,sig,log=T)
      #tmp = dmvnorm(y,x%*%beta,sig,log=T)
      #print(tmp)
      ll = ll+tmp
    }
    return(-ll)
}


pars_male=c("mu"=0.5,"sex_male"=-0.1,"alpha_male"=0,"sigma_sub_day^2"=0,"sigma_s^2"=0,"sigma^2"=0)
pars_female=c("mu"=0.5,"sex_male"=-0.1,"alpha_female"=0,"sigma_sub_day^2"=0,"sigma_s^2"=0,"sigma^2"=0)

nll2(pars)

# Fitting male alpha, female=1
opt2<-nlminb(pars_male,nll2,alpha_sex="male")
out_par = opt2$par
c(out_par[1:2],exp(out_par[3:6]))
exp(out_par[3])*exp(out_par[4:6])

# Fitting female alpha, male=1
opt2<-nlminb(pars_female,nll2,alpha_sex="female")
out_par = opt2$par
c(out_par[1:2],exp(out_par[3:6]))
exp(out_par[3])*exp(out_par[4:6])


```

